Supplement to: Optimal management of a stochastically varying population when policy adjustment is costly

### 2015-01-15 manuscripts/manuscript.Rmd
working on edits from paul and mike

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTKEEB

@@

abstract: |
          Ecological systems are dynamic and policies to manage them need to
          respond to that variation. However, policy adjustments will sometimes
          be costly, which means that fine-tuning a policy to track variability in
          the environment very tightly will only sometimes be worthwhile. We use a
          classic fisheries management question -- how to manage a stochastically
          varying population using annually varying quotas in order to maximize
          profit  -- to examine how costs of policy adjustment change optimal
          management recommendations. Costs of policy adjustment (here changes in
          fishing quotas through time) could take different forms. For example,
          these costs may respond to the size of the change being implemented or
          there could be a fixed cost any time a quota change is made. We show
          how different forms of policy costs have contrasting implications for
          optimal policies. Though it is frequently assumed that costs to adjusting
          policies will dampen variation in the policy, we show that certain cost
          structures can actually increase variation through time. We further show
          that failing to account for adjustment costs has a consistently worse
          economic impact than would assuming these costs are present when they are not.

@@

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space [@Durrett1994]. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to reflect environmental variation more closely
should be traded off against potential costs associated with the more
interventionist approach to management this would require.

@@

To illustrate this tradeoff, we use an example from fisheries
management. Figure 1, top-left panel, shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (_Thunnus thynnus_, henceforth bluefin) from
a recent stock assessment [@ICCAT20XX].  The bottom-left panel 
shows the catch quota (Total Allowable Catch or TAC, before being adjusted
for overages or underages in the catch of each country in the previous year)
for the stock as set by the relevant
management agency [@ICCAT20XX]. Despite the estimated population size
declining by XX% between YYYY and ZZZZ, the quota was not changed during
this period. Instead, the quota only changed occasionally and in between
times was left unaltered. Fishery management decisions regarding this
species can be highly contentious (@Safina1998, @Sissenwine1998,
@Porch2005). Moreover, the stock is fished by fleets from many nations
with quotas being set by a multilaterial management agency through a
process of negotiation. As such, we might reasonably anticipate that
for this species there could be substantial transaction costs involved
in reaching agreement over any quota change, which could contribute to
the observed quota stability. Such patterns of variable stocks managed
under far less variable quotas are extremely common; for instance, 
in the management of south Atlantic snowy grouper (_Epinephelus niveatus_) 
shown in the right-hand panels [@SEDAR2013]. More generally, reviews by @Biais1995 and
@Patterson2007a document many cases where changes in catch
quotas that a management agency set were more modest than changes that
would be recommended just by considering variations in stock abundance.

@@

With his formulation, @Reed1979 showed that a constant escapement policy could be
optimal under certain conditions. Such a policy involves choosing annual
quotas that are perfectly responsive to recruitment variation in a fish
stock. In poor recruitment years, the quota is set to zero and no fishing
is allowed. But any time there is a good recruitment pulse, a quota is
set that allows the fishery to exactly compensate through harvesting,
thereby maintaining the optimal escapement level. However, in that
analysis, Reed did not account for any costs of policy adjustment, which
for such a responsive management strategy potentially could be large.

As the examples in Figure 1 make clear, management policies will rarely be
as responsive as this constant escapement policy assumes. In this paper,
we explore conditions under which this lack of responsiveness may be a
rational response on the part of managers. Specifically, we consider a
case where managers seek to balance the benefits in terms of increased
profits from fishing from more finely tracking recruitment variations
with the growing costs associated with adjusting policies quickly to do
so. The policy adjustment costs involved could reflect pure administrative
transaction costs or preferences held by fishermen, fish processing plants
or other stakeholders, for less variable quotas. In seeking to account
for these policy adjustment costs, we recognize that we do not know just
what functional form they should take and that it will require a lot of
empirical work to estimate that. Therefore, we scope three candidate
functional forms that represent different assumptions about how these
costs operate to determine whether the results we obtain are sensitive
to such differences.

@@

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$.

To simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the intrinsic rate of increase of
the fish population is 1 and the equilibrium biomass for the equivalent
deterministic model without harvesting is 10; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn \eqref{eq:state_equation}. To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.
In the supplement we illustrate that the qualitative patterns observed
here are not sensitive to the specific choices for these parameter values.

Taken together this objective function \eqref{eq:objective} and the state
equation \eqref{eq:state_equation} define a stochastic dynamic programming
problem that we solve using backwards recursion via Bellman's equation.
We denote the resulting state-dependent, optimal control as $h_0^{\ast}$.
We solve this problem on a finite time horizon of $T =$ `r OptTime` using
value iteration [@Mangel1988; @Clark2000].

Note that this problem is much larger computationally than the classic 
formulation of this stochastic dynamic programming problem for a single stock.
Whereas the classic problem considers each of $Q$ possible discrete quotas
for $S$ possible stock values at each time $t$, for a search space of size $S \times Q$
per timestep;  this formulation must also consider all possible values of quota
in the _previous_ time step: a total of $S \times Q \times Q$ configurations.

@@

where indicator function $\mathbf{I}(h_t, h_{t-1})$ takes the value 1 if
$h_t=h_{t-1}$ and zero otherwise. For each $\Pi_i$, we can then define a
new objective function $NPV_i$ similar to that in Eqn. \eqref{eq:objective}.

We note that the fixed cost has conceptual analogues to the set-up cost of
@Reed1974 and @Spulber1982 but here the fee is associated with a change
in harvesting ($h_t\neq h_{t-1}$) and not with harvesting *per se*
($h_t>0$) as in those studies.

Taken together with the state equation Eqn. \eqref{eq:state_equation}, each
of these new objective functions defines a different stochastic dynamic
programming problem. Again, we solve them numerically using backwards
recursion. To include costs of policy adjustment, we expand the state
space to include both the current stock size and the management action
taken on the previous time step $(N_t,h_{t-1})$. For each new objective
function $\max NPV_i$, we denote the corresponding optimal control
policy $h_i^*$.

@@

<!-- State conclusion of this supplement here! -->

@@

Before we can compare the policy dynamics resulting from each of these
functional froms, we must be careful to ensure that we are only varying
the _structure_ of the adjustment penalty and not the overall magnitude.
At the extremes, a large enough adjustment cost can ensure a constant 
policy, regardless of the structure; while a small enough one can be ignored
regardless.  Defining this calibration correctly is thus key to the entire
analysis.  Fortunately, the optimal control framework provides a completely
natural way to make this comparison: in terms of the _economic value_
(measured by $NPV$) of the stock in each case.  We will calibrate this relative
to economic value of the stock when adjustments to the policy are free ($NPV_0$).

Each policy adjustment cost function is characterized in terms of
a cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal NPV
fishery.

Figure 2 shows the calibration graphically. Each curve plots the change
in maximum expected $NPV$ for a given penalty function as the penalty cost
parameter is increased. In each case, the figure shows maximum expected
NPV with policy adjustment costs as a proportion of the maximum expected
NPV available in the basic problem $NPV_0({\bf h_0^*})$ without policy
adjustment costs. The $c_i$ values are given in Table 1.

```{r table_1, results="asis"}
pandoc.table(print_npv)
```

@@

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients that
induce the same reduction in maximum expected NPV. For example, the
dashed vertical lines in Figure 2 map the needed cost coefficient $c_i$,
for each penalty function, such that the fishery is worth `r 100*(1-reduction)`% of its
unconstrained value when optimally managed. We focus our examination of the models on how
the type and severity of policy adjustment costs affects the optimal management strategy.

@@

ggplot(dt3, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Harvest Dynamics")  +
  scale_color_discrete(labels=c("With Penalty", "Reed Solution"), name="") +
  scale_linetype_discrete(labels=c("With Penalty", "Reed Solution"), name="")

@@

Figure 3 illustrates how the different forms of adjustment cost can impact
the optimal harvest policy. Corresponding stock sizes can be seen in the
supplement.  Each panel is generated against the same sequence of 
environmental variability so that they can be compared directly.  The harvest policy chosen
shows a systematic deviations from the penalty free optimum, depending on
the nature of the cost.  In each case, the optimal solution without any
adjustment cost is shown with the policy induced by optimization
under the given adjustment penalty (equivalent to a `r 100*reduction`%
reduction in maximum expected NPV) overlaid in blue.

The first panel shows a typical pattern resulting from linear adjustment costs ($\Pi_1$).
The optimal policy tends to avoid very small policy adjustments, resulting in
periods of a constant policy followed by sudden bursts of adjustment.
This results in a relatively step-like policy pattern. In contrast, the
second formulation (quadratic costs, $\Pi_2$) disproportionately penalizes large policy
adjustments. The corresponding optimal policy is typified by the middle panel, responding
to each of the fluctuations in stock, made by the cost-free policy, but with smaller 
magnitude response than the equivalent cost-free optimal solution. This results
in a smoother $h_t$ curve, one that undershoots the larger oscillations
seen in the cost-free optimum in favor of a policy that changes
incrementally each year. Finally, the optimal policy for the third,
'fixed fee' formulation ($\Pi_3$) only makes large adjustments, as one
might expect, because the magnitude of the adjustment made is not
reflected in the resulting penalty.

These patterns are consistent across stochastic replicates over a range
of penalty magnitudes. The comparisons shown in Figure 3 are for one 
realization (i.e. a particular sequence of random number draws representing
environmental variability) and are made for one particular magnitude of 
policy adjustment costs (Calibrated to equal to `r reduction *100`% of the 
maximum expected net present valuein the base case ($NPV_0({\bf h_0^*})$,
as discussed in Figure 2, permitting comparisons between different forms).
To demonstrate this, we solved for the optimal policy under each of the
three adjustment cost scenarios for `r length(c0)` different penalty 
coefficients. For each resulting policy, we then simulated 50 stochastic
replicates of the stock dynamics managed under that policy.  To summarize
the patterns shown in Figure 3 and characterize the impact of increasingly
large fraction of the economic value (NPV) being consumed by policy
adjustment costs, we examine the response of several summary statistics
to increasing adjustment penalties (as a fraction of the adjustment-free value)
in Figure 4.  We show the impact of policy adjustment costs on the
variance & autocorrelation of the harvests through time.

@@

While the optimization guarentees that on average, the policy in the first simulation
should outperform that of the second, it does not tell us by how much.  These simulations
thus give us a lense to examine the magnitude of this effect.

We provide 100 replicates of each kind for each of the three forms
$\Pi_{1-3}$ (in which each ensemble of 100 replicates uses the same sequence of random seeds).
For each, we record (a) the dockside profits derived from fishing as well
as (b) any fees paid to make adjustments to the policy.  In the first kind of 
simulation, dockyard profits are almost less than in the second for a given stochastic
replicate, since the latter optimizes the expected dockside NPV.  However, the costs
paid for adjustment are almost always less in the first scenario as well, since
this solution explicitly tries to avoid higher adjustment costs.  We visualize this
pattern in the distributions shown in Figure 5 (focusing on $\Pi_1$ for simplicity,
as the pattern in $\Pi_2$ and $\Pi_3$ is similar, see Supplementary Material.)

Figure 5 shows that the difference in dockside revenue (labelled "profits" in the figure) 
between these two scenarios 
(green vs purple distributions) is very small, while a larger difference exists between
the costs paid for making policy adjustments (red vs blue). This is predicted by the optimization: 
the first scenario must outperform (or at least equal) the NPV of the second scenario
in aggregate, and as it cannot do so through higher profits it must do so through lower
costs.  Interestingly, for this set of parameters very little is lost in dockside revenue.
This suggests that the impact of assuming costs are present when they are not is rather small:
since the dockside revenue is none-the-less near optimal.  This also suggests that the 
converse error: ignoring costs when they are present, is more severe.

@@

ggplot(filter(table1_long, reduction == 0.25),
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figure6.csv")
```

@@

Figure 6 shows that across each penalty form $\Pi{1-3}$, the economic impact
of ignoring adjustment costs when they are present is higher than accounting
for them when they are absent. This holds even after accounting for effect 
described above: the red bars are consistently well below (`r (1-reduction)`),
while the blue bars are barely below 1.
We also see that the impacts on net present value of the fishery of
incorrectly assuming or ignoring policy adjustment costs are most severe
when these costs are assumed to scale quadratically with the size of the
policy change being implemented ($\Pi_2$).
In the Supplementary Materials we show that pattern holds across different
severity of adjustment costs (robust to choice of $c_i$), though the larger
the adjustment cost, the greater the difference between these two errors.

Note that this analysis does not consider the possibility that an adjustment cost
is present, but of a different functional form than the policy assumes.  Not surprisingly,
the consequences of managing under the wrong functional form can be as or more severe than
simply ignoring adjustment costs in the optimization all-together.  We illustrate several
of these mismatch errors in the Supplementary Material.

@@

This work was assisted through participation in "Pretty Darn Good Control" Investigative Workshop at the National Institute for Mathematical and Biological Synthesis, sponsored by the National Science Foundation through NSF Award #DBI-1300426, with additional support from The University of Tennessee, Knoxville.  The authors acknowledge helpful discussions and input from working group co-organizers Megan Donahue and Carl Toews, and participants Marie-Josee Fortin, Dan Ryan, Frank Doyle, Claire Paris, Iadine Chades and Mandy Karnauskas.
We also acknowledge the following support: NSF Grant DBI-1306697 (CB), ...

>>>>

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul R. Armsworth 
    affiliation: UTKEEB

@@

abstract: |
          Ecological systems are dynamic and policies to manage them need to
          respond to that variation. However, policy adjustments will sometimes
          be costly, which means that fine-tuning a policy to track variability in
          the environment very tightly will only sometimes be worthwhile. We use a
          classic fisheries management question -- how to manage a stochastically
          varying population using annually varying quotas in order to maximize
          profit  -- to examine how costs of policy adjustment change optimal
          management recommendations. Costs of policy adjustment (here changes in
          fishing quotas through time) could take different forms. For example,
          these costs may respond to the size of the change being implemented, or
          there could be a fixed cost any time a quota change is made. We show
          how different forms of policy costs have contrasting implications for
          optimal policies. Though it is frequently assumed that costs to adjusting
          policies will dampen variation in the policy, we show that certain cost
          structures can actually increase variation through time. We further show
          that failing to account for adjustment costs has a consistently worse
          economic impact than would assuming these costs are present when they are not.

@@

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space [@Durrett1994]. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to more closely reflect environmental variation
should be traded off against potential costs associated with the more
responsive approach to management this would require.

@@

To illustrate this tradeoff, we use an example from fisheries
management. Figure 1, top-left panel, shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (_Thunnus thynnus_, henceforth bluefin) from
a recent stock assessment [@ICCAT20XX].  The bottom-left panel 
shows the catch quota (Total Allowable Catch or TAC, before being adjusted
for overages or underages in the catch of each country in the previous year)
for the stock as set by the relevant
management agency [@ICCAT20XX]. Despite the estimated population size
declining by XX% between YYYY and ZZZZ, the quota was not changed during
this period. Instead, the quota only changed occasionally and in between
times was left unaltered. Fishery management decisions regarding this
species can be highly contentious (@Safina1998, @Sissenwine1998,
@Porch2005). Moreover, the stock is fished by fleets from many nations
with quotas being set by a multilaterial management agency through a
process of negotiation. As such, we might reasonably anticipate that
for this species there could be substantial transaction costs involved
in reaching agreement over any quota change, or constraints on the frequency
with which the those changes could occur.  Either could contribute to
the observed quota stability. Such patterns of variable stocks managed
under far less variable quotas are extremely common; for instance, 
in the management of south Atlantic snowy grouper (_Epinephelus niveatus_) 
shown in the right-hand panels [@SEDAR2013]. More generally, reviews by @Biais1995 and
@Patterson2007a document many cases where changes in catch
quotas that a management agency set were more modest than changes that
would be recommended just by considering variations in stock abundance.

@@

With his formulation, @Reed1979 showed that a constant escapement policy could be
optimal under certain conditions. Such a policy involves choosing annual
quotas that are perfectly responsive to recruitment variation in a fish
stock. In poor recruitment years, the quota is set to zero and no fishing
is allowed. Any time a recruitment pulse exceeds the escapement threshold, a quota is
set that allows the fishery to exactly compensate through harvesting,
thereby maintaining the optimal escapement level. However, in that
analysis, Reed did not account for any costs of policy adjustment, which
for such a responsive management strategy potentially could be large.

As the examples in Figure 1 make clear, management policies will rarely be
as responsive as this constant escapement policy assumes. In this paper,
we explore conditions under which this lack of responsiveness may be a
rational response on the part of managers. Specifically, we consider a
case where managers seek to balance the benefits in terms of increased
profits from fishing from more finely tracking recruitment variations
with the growing costs associated with adjusting policies frequently to do
so. The policy adjustment costs involved could reflect pure administrative
transaction costs or preferences held by fishermen, fish processing plants
or other stakeholders, for less variable quotas. In seeking to account
for these policy adjustment costs, we recognize that we do not know just
what functional form they should take and that it will require a lot of
empirical work to estimate that. Therefore, we scope three candidate
functional forms that represent qualitatively different assumptions about how these
costs operate to determine whether the results we obtain are sensitive
to such differences.

@@

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing per unit effort and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$.

To simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the intrinsic rate of increase of
the fish population is 1 and the equilibrium biomass for the equivalent
deterministic model without harvesting is `r K`; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn \eqref{eq:state_equation}). To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.
In the supplement we illustrate that the qualitative patterns observed
here are not sensitive to the specific choices of these parameter values.

Taken together this objective function \eqref{eq:objective} and the state
equation \eqref{eq:state_equation} define a stochastic dynamic programming
problem that we solve using backwards recursion via Bellman's equation.
We denote the resulting state-dependent, optimal control as $\mathbf{h}_0^{\ast}$.
We solve this problem on a finite time horizon of $T =$ `r OptTime` using
value iteration [@Mangel1988; @Clark2000]. (Longer horizon times have no
qualitative difference but stretch the plots making it more difficult
to visualize the fluctuations.)

@@

where indicator function $\mathbf{I}(h_t, h_{t-1})$ takes the value 1 if
$h_t=h_{t-1}$ and zero otherwise. For each $\Pi_i$, we can then define a
new objective function $NPV_i$ similar to that in Eqn. \eqref{eq:objective}.
We note that the fixed cost has conceptual analogues to the set-up cost of
@Reed1974 and @Spulber1982 but here the fee is associated with a change
in harvesting ($h_t\neq h_{t-1}$) and not with harvesting *per se*
($h_t>0$) as in those studies.

Taken together with the state equation Eqn. \eqref{eq:state_equation}, each
of these new objective functions defines a different stochastic dynamic
programming problem. Again, we solve them numerically using backwards
recursion. To include costs of policy adjustment, we expand the state
space to include both the current stock size and the management action
taken on the previous time step $(N_t,h_{t-1})$. For each new objective
function $\max NPV_i$, we denote the corresponding optimal control
policy as the vector $\mathbf{h}_i^*$.

Note that this problem is much larger computationally than the classic 
formulation of this stochastic dynamic programming problem for a single stock.
Whereas the classic problem considers each of $Q$ possible discrete quotas
for $S$ possible stock values at each time $t$, for a search space of size $S \times Q$
per timestep;  this formulation must also consider all possible values of quota
in the _previous_ time step: since the costs that will follow depend on whether
and by how much the harvest policy will change. This creates a total of
$S \times Q \times Q$ configurations.

@@

@@

Each policy adjustment cost function is characterized in terms of a
cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect
of each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal
fishery $NPV$.  Fortunately, the optimal control framework provides a
completely natural way to make this comparison: in terms of the _economic
value_ (measured by $NPV$) of the stock in each case.  We will calibrate
our realized economic value under each policy relative to economic value
of the stock when adjustments to the policy are free ($NPV_0$).

Figure 2 shows the calibration graphically. Each curve plots the change in
maximum expected $NPV$ for a given cost structure as the $c_i$ coefficient
is increased. In each case, the figure shows maximum expected
NPV with policy adjustment costs as a proportion of the maximum expected
NPV available in the basic problem $NPV_0({\bf h_0^*})$ without policy
adjustment costs.

@@

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients
that induce the same reduction in maximum expected NPV. For example,
the dashed vertical line in Figure 2 maps the needed cost coefficient
$c_i$, for each penalty function, such that the fishery is worth `r
100*(1-reduction)`% of its unconstrained value when optimally managed.

@@

ggplot(dt3, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Harvest Dynamics")  +
  scale_color_discrete(labels=c("Adjustment penalty", "No penalty"), name="") +
  scale_linetype_discrete(labels=c("Adjustment penalty", "No penalty"), name="")

@@

Figure 3 illustrates how the different forms of adjustment cost can
impact the dynamics of the optimal harvest policy. Corresponding stock
sizes can be seen in the supplement.  Each panel is generated against the
same sequence of environmental variability so that they can be compared
directly.  The harvest policy chosen shows a systematic deviations from
the adjustment cost free optimum, depending on the structure of the cost function.  In each
case, the optimal solution without any adjustment cost is shown with
the policy induced by optimization under the given cost structure 
(equivalent to a `r 100*reduction`% reduction in maximum expected NPV)
overlaid in blue.

The first panel shows a typical pattern resulting from linear adjustment
costs ($\Pi_1$).  The optimal policy tends to avoid very small policy
adjustments, resulting in periods of a constant policy followed by
sudden bursts of adjustment.  This results in a relatively step-like
policy pattern. In contrast, the second formulation (quadratic costs,
$\Pi_2$) disproportionately penalizes large policy adjustments. The
corresponding optimal policy is typified by the middle panel, responding
to each of the fluctuations in stock, made by the cost-free policy, but
with smaller magnitude response than the equivalent cost-free optimal
solution. This results in a smoother $h_t$ curve, one that undershoots
the larger oscillations seen in the cost-free optimum in favor of a policy
that changes incrementally each year. Finally, the optimal policy for the
third, 'fixed fee' formulation ($\Pi_3$) only makes large adjustments,
as one might expect, because the magnitude of the adjustment made is
not reflected in the resulting cost.

These patterns are consistent across stochastic replicates over a
range of penalty magnitudes. The comparisons shown in Figure 3 are
for one realization (i.e. a particular sequence of random number draws
representing environmental variability) and are made for one particular
magnitude of policy adjustment costs, which have been calibrated to equal
to `r reduction *100`% of $NPV_0({\bf h_0^*})$.  To demonstrate this,
we solved for the optimal policy under each of the three adjustment
cost scenarios for `r length(c2)` different $c_i$ coefficients. For each
resulting policy, we then simulated 50 stochastic replicates of the stock
dynamics managed under that policy.  To summarize the patterns shown
in Figure 3 and characterize the impact of increasingly large fraction
of the economic value (NPV) being consumed by policy adjustment costs,
we examine the response of several summary statistics to increasing
adjustment penalties (as a fraction of the adjustment-free value) in
Figure 4.  We show the impact of policy adjustment costs on the variance &
autocorrelation of the harvests through time.

@@

While the optimization ensures that on average, the policy in the first
simulation should outperform that of the second, it does not tell us by
how much.  These simulations thus allow us to quantify this effect.

We provide 100 replicates of each kind of scenario for each of the
three forms $\Pi_1$, $\Pi_2$ & $\Pi_3$.  For each, we record (a)
the dockside profits derived from fishing as well as (b) any fees paid
to make adjustments to the policy.  In the first kind of simulation,
dockside revenues are almost less than in the second for a given
stochastic replicate, since the latter optimizes the expected NPV.
However, the costs paid for adjustment are almost always less in the
first scenario as well, since this solution explicitly tries to avoid
higher adjustment costs.  We visualize this pattern in the distributions
shown in Figure 5 (focusing on $\Pi_1$ for simplicity, as the pattern
in $\Pi_2$ and $\Pi_3$ is similar, see Supplementary Material.)

Figure 5 shows that the difference in dockside revenue (labeled "profits"
in the figure) between these two scenarios (green vs purple distributions)
is very small, while a larger difference exists between the costs paid
for making policy adjustments (red vs blue). This is predicted by the
optimization: the first scenario must outperform (or at least equal)
the NPV of the second scenario in aggregate, and as it cannot do so
through higher profits it must do so through lower costs.  Interestingly,
for this set of parameters very little is lost in dockside revenue.
This suggests that the impact of assuming costs are present when they
are not is rather small: since the dockside revenue is none-the-less
near optimal.  This also suggests that the converse error: ignoring
costs when they are present, is more severe.

@@

ggplot(filter(table1_long, reduction == 0.25),
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function") + 
  ylab(expression(NPV[0]))
write.csv(table1_long, "components/data/figure6.csv")
```

@@

Figure 6 shows that across each penalty form $\Pi_1$, $\Pi_2$ & $\Pi_3$,
the economic impact of ignoring adjustment costs when they are present
is higher than accounting for them when they are absent. This holds even
after accounting for effect described above: the red bars are consistently
well below the `r (1-reduction)*100`% of $NPV_0$, while the blue bars
are barely below $NPV_0$ (a value of 1 on the vertical axis).  We also
see that the impacts on net present value of the fishery of incorrectly
assuming or ignoring policy adjustment costs are most severe when these
costs are assumed to scale quadratically with the size of the policy
change being implemented ($\Pi_2$).  In the Supplementary Materials we
show that pattern holds across different severity of adjustment costs
(robust to choice of $c_i$), though the larger the adjustment cost,
the greater the difference between these two errors.

Note that this analysis does not consider the possibility that an
adjustment cost is present, but of a different functional form than the
policy assumes.  The consequences of managing under the wrong functional
form can be as or more severe than simply ignoring adjustment costs in
the optimization altogether.  We illustrate several of these mismatch
errors in the Supplementary Material.

@@

This work was assisted through participation in "Pretty Darn Good Control"
Investigative Workshop at the National Institute for Mathematical and
Biological Synthesis, sponsored by the National Science Foundation through
NSF Award #DBI-1300426, with additional support from The University of
Tennessee, Knoxville.  The authors acknowledge helpful discussions and
input from working group co-organizers Megan Donahue and Carl Toews, and
participants Marie-Josee Fortin, Dan Ryan, Frank Doyle, Claire Paris,
Iadine Chades and Mandy Karnauskas.  We also acknowledge the following
support: NSF Grant DBI-1306697 (CB), ARC DECRA (DE130100572) and The
ARC Centre of Excellence for Environmental Decisions (MB), ...

### 2015-01-15 manuscripts/supplement.Rmd
add nimbios ack
supplemental data files
working on edits from paul and mike

---
title: "Supplement to: Optimal management of a stochastically varying population when policy adjustment is costly"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTKEEB

@@

@@

@@

Recall that in our model definition, stock sizes are measured before
harvest ($h_t$ being determined after observing stock size $x_t$) and
thus show less influence of the policy cost than the harvest itself.

@@

write.csv(ex, "components/data/figure_s3.csv")
```

@@

```{r quadraticcosts, include=FALSE, message=FALSE}

source("components/quad_costs.R")
```

>>>>

---
title: "Supplement to: Optimal management of a stochastically varying population when policy adjustment is costly"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul R. Armsworth 
    affiliation: UTKEEB

@@

write.csv(dt_s, "components/data/figureS1.csv")
```

@@

write.csv(stock_stats, "components/data/figureS2.csv")

@@

Recall that in our model definition, stock sizes are measured before
harvest ($h_t$ being determined after observing stock size $N_t$) and
thus show less influence of the policy cost than the harvest itself.

@@

write.csv(ex, "components/data/figureS3.csv")
```

@@

```{r quadraticcosts, include=FALSE, message=FALSE}
source("components/quad_costs.R")
```
